Linear response theory for core set identification

### 2013-06-04 writedown/linear.response/paper.tex
rewrite the result part, so that it is consistent with the updated theoy language

\newcommand{\mh}{\mathcal H}
\newcommand{\ml}{\mathcal L}
\newcommand{\mt}{\mathcal T}
\newcommand{\mo}{\mathcal O}
\newcommand{\mi}{\mathcal I}
\newcommand{\mc}{\mathcal C}
\newcommand{\proj}{\mathit\Pi}
\newcommand{\fwg}{{\mathcal A}}
\newcommand{\bwg}{{\mathcal B}}
\newcommand{\bsigma}{\boldsymbol\sigma}

@@

\title{Linear response theory for core set identification}
\author{Han Wang}
% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

\section{Theoretical background}
\subsection{Classical linear response theory}

In this section, we want to firstly give a brief description of the classical linear response
theory, for details see Ref.~\cite{tuckeman2010statistical} for
example.
In this paper we want to study the non-equilibrium system driven by some external force.
More strictly, we investigate the system that is governed by the
following Langevin equation:
\begin{align}\label{eqn:langevin-1}
  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2}
  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
  + \vect D(\vect q,\vect p) F_e(t)
  - \gamma\vect p
  + \sigma\dot{\vect W}.
Here $\vect D$ is the external driving force applied to the system, and
$F_e(t)$ is the strength of the driving as a function of time.
We always assume that the magnitude of $\vect D$ is controllable, and
the magnitude of $F_e(t)$ is small and of order $\mo (\varepsilon)$, $\varepsilon \ll 1$.
The phase space incompressibility condition is assumed:
  \nabla_{\vect p}\cdot\vect D = 0
Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
q, \vect p\}$) is subject to the Kolmogorov forward equation:
\begin{align}\label{eqn:orig-forward}
  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
  \fwg^\beta(t) =
  \frac{\sigma^2}2\Delta_{\vect p}
  - \Big(
  \nabla_{\vect p}\mh + \vect C F_e(t)
  \Big)\cdot\nabla_{\vect q}
  - \Big(
  -\nabla_{\vect q}\mh +
  \vect D F_e(t) - \gamma\vect p
  \Big)\cdot\nabla_{\vect p}
  + 3N\gamma,
and $\beta$ is the inverse temperature.
The fluctuation-dissipation relation is always true:
$\beta = 2\gamma / \sigma^2$.
The equilibrium and non-equilibrium system
is governed by
the standard Langevin equation, i.e. with
vanishing driving force $F_e(t) = 0$ in
Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.
We assume that the driving to the
system is so small that both the phase space distribution and the
infiniesimal generator can be viewed as perturbation from the initial
distribution and the un-driven generator
($\vect D = 0$),
respectively. We the expansion with respect to small
variable $\varepsilon$:
\begin{align}\label{eqn:f-expan}
  f(\vect x, t) &= f_0(\vect x, t) + \varepsilon f_1(\vect x, t)
  +\varepsilon^2 f_2(\vect x, t) + \mo (\varepsilon^3),
and 
\begin{align}\label{eqn:A-expan}
  \fwg^\beta(t) = \fwg^\beta_0 + \varepsilon\fwg_1(t),
where
  \fwg_0^\beta =&
  \frac{\sigma^2}2\Delta_{\vect p}
  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
  - \big(
  -\nabla_{\vect q}\mh - \gamma\vect p
  \big)\cdot\nabla_{\vect p}
  + 3N\gamma,\\
  \fwg_1(t) =&
  - \frac{F_e(t)}{\varepsilon} \big(
  \vect C \cdot\nabla_{\vect q}
  +
  \vect D \cdot\nabla_{\vect p}
  \big).
Notice $F_e(t) \sim \mo(\varepsilon)$, so $\Vert\fwg_1(t)\Vert\sim\mo(1)$.
The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$,
which is nothing but the Boltzmann distribution

@@

  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]

@@

  [\,\fwg^\beta_0 + \varepsilon\fwg_1(t)\,]
  [f_0(\vect x, t) + \varepsilon  f_1(\vect x, t) + \varepsilon^2f_2(\vect x, t)]

@@

Matching this equation at different orders of $\varepsilon$, the zeroth

@@

which is the un-driven equation. Matching the first and the second order gives:

@@

  -\frac{F_e(s)}{\varepsilon}\beta

@@

  -\vect C\cdot\nabla_{\vect q}\mh 

@@

  -\frac\beta\varepsilon

@@

  - \beta

@@

  + \mo(\varepsilon^2)

@@

For any time-dependent observation, the linear response approximation
would be:

@@

  [\,f_{\inv}^\beta(\vect x) + \varepsilon f_1(\vect x, t)] \\\nonumber

@@

  \beta
  \int \dd \vect x\:
  O(\vect x)
  \int_0^t\dd s\,F_e(s)
  e^{(t-s)\fwg_0^\beta}
  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
  \\\nonumber
  \beta
  F_e(s)
  O(\vect x)\,
  e^{(t-s)\fwg_0^\beta}
  \big[j(\vect x)f^\beta_{\inv}(\vect x)\big]
  \\\nonumber
  \beta
  F_e(s)
  e^{(t-s)\bwg^\beta_0}
  O(\vect x)\\ \label{eqn:eqi-pert-1}
  \beta

@@

  O(\vect x)

@@

Where $\bwg^\beta_0$ is the infiniesimal generator of the \emph{un-driven} 
Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$.
We have

@@

where $\vect X^\beta_t$ is the trajectory of the un-driven Langevin equation

@@

Therefore, we have

@@

@@

  \beta

@@

  \mathbb E_{\vect x} [O(\vect X^\beta_s)]
or equivalently:
\begin{align}\label{eqn:lr}

@@

  \beta

@@

  j(0)\,
  \mathbb E_{\vect x} [O(\vect X^\beta_s)]

@@

  correlation} between the observation $O$ at time $s$ and the
dissipative flux $j$.

@@

The classical response theory mainly has the following two limitations:
Firstly, one has to start from an equilibrium simulation. That means if the
external driving is no longer small ($\varepsilon$ is relatively big), the
first order (or linear) response~\eqref{eqn:lr} may not be a good approximation to
the true non-equilibrium process any more.
One possible way to improve the accuracy is to introduce the second
order response term, i.e.~\eqref{eqn:f2-0}.
Here one encounters the second limitation of the classical
response theory: it does not provide a frame work, under which
the second and even higher order responses can be easily derived.
The limitations of the classical responses theory motivate us to
develop a more general frame work that allows one starting from any
non-equilibrium process (arbitrarily driven), and calculating from it as many orders of
responses as necessary.

In the language of path integral, the time-dependent average $O( t)$
can be written as
  O( t) =
  \langle O[\vect X_s]\,\rangle_{0,t} = 
  \int\dd \vect u\, f(\vect u, 0) 
  \int_{\mc\{\vect u,0; t\}}
  O[\vect X_s]\, \dd\mathcal P[\vect X_s]  
where $O[\vect X_s]$ is the observable of the trajectories, which is a generalization of the
observable that only depends on the configuration of the system.
$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories in time interval $[0,t]$.
$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution.
$\mc\{\vect u,0; t\}$ is the set of all continuous trajectories starting from point
$\vect u$ at time $0$, and ending at some position at time $t$.
$\mathcal P[\vect X_s]$ is the
probability measure of the trajectories generated by driven Langevin dynamics~\eqref{eqn:langevin-1}--\eqref{eqn:langevin-2},
where the intensity $F_e$ need not to be 0.
Now we want to consider a system that is a small perturbation of the previous system (we now call it ``reference system'')
in the sense that it has a similar driven force: $F_e + \Delta F_e$, where $\Delta F_e\sim \mo(\varepsilon)$ is small.
The probability measure of the trajectories of this perturbed system is denoted by
$\tilde{\mathcal P}[\vect X_s]$.
The time-dependent average of observable of the system $F_e + \Delta F_e$ can be written down in terms of
the probability measure of the reference system:
  \tilde O(t) =
  \int \dd\vect u f(\vect x,0)
  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
  \frac{\dd\tilde{\mathcal P}[\vect X_s]}{\dd\mathcal P[\vect X_s]}
  \, \dd\mathcal P[\vect X_s]    
by using the Girsanov formula, 
\begin{align}\label{eqn:girsanov}
  \frac
  {\dd\tilde{\mathcal P}[\vect X_s]}
  {\dd\mathcal P[\vect X_s]}  
  \exp
  \big\{{G[\vect X_s] - \frac12 H[\vect X_s]}
  \big\}
where the short notations are defined by:
  G[\vect X_s]
  &= \frac{1}\sigma\int_0^t
  \Delta F_e(s)\,\vect D(\vect q_s,\vect p_s)\cdot\dd\vect W_s \\
  H[\vect X_s]
  \frac 1{2\sigma^2}\int_0^t
  \Delta F_e^2(s)\,\vect D^2(\vect q_s,\vect p_s)\,\dd s
If the difference $\Delta F_e(t)$ is of order $\mo (\varepsilon)$, then $G[\vect X_s] \sim \varepsilon$ is
the first order term, and $H[\vect X_s] \sim \varepsilon^2$  is the second order term.
Then with the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}, we have
\begin{align}\label{eqn:neq-response}
  \tilde O(t) -
  O(t)
  \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
  +
  \frac12 \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
  +
  \mo (\varepsilon^3)
Here are some remarks of the main result Eq.~\eqref{eqn:neq-response}:
\begin{itemize}
\item The first term on the right-hand-side of Eq.~\eqref{eqn:neq-response} is the linear
  response of the reference non-equilibrium process (driven by $F_e$). The second term is the second order
  response. One can easily write down higher orders of response by the Taylor expansion
  of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
\item When the observable $O[\vect X_s]$ only depends on the end point of the trajectory, i.e.
  $O[\vect X_s] = O(\vect X_t)$, then $\big\langle O(\vect X_t)\cdot G[\vect X_s]\big\rangle_{0,t}$
  gives the well-know time correlation form of the linear response.
\item If one wants to calculate a large amount of different perturbations, then apply one by one
  Eq.~\eqref{eqn:neq-response} is not convenient, because one has to calculate the responses
  again when the function form of $\Delta F_e$ is not a linear combination of previously calculated perturbations.
  If it is possible to
  express different $\Delta F_e$ by a group of base functions, then the responses are only need
  to calculated on the bases. Then with a linear conbination of the responses on bases, one can derive
  the responses of the whole group of perturbations. 
\item In the prove, we assume that the reference and the perturbed non-equilibrium processes
  have the same initial distribution (both of them may be different from the equilibrium distribution),
  but this is not a necessary. One can easily
  applies similar perturbation approach to the initial distribution as  to the trajectory ensemble.
  To keep the story simple, we do not develop the theory from this aspect.
\end{itemize}

\section{Numerical tests}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
  \caption{The double-well potential with tilting perturbation.}
  \label{fig:tmp1}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str2-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 2$.
    The 1st row: the direct non-equilibrium
    simulation. The 2nd row: the results of linear response
    formula~\eqref{eqn:core-identify-approx-2}.
    The 3rd row: the results of response
    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
    the equilibrium state.
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str4-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $F_e^{\textrm{max}} = 4$.
    The 1st row: the direct non-equilibrium
    simulation. The 2nd row: the results of linear response
    formula~\eqref{eqn:core-identify-approx-2}.
    The 3rd row: the results of response
    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
    The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
    the equilibrium state.
    The 4th row: the results of response
    formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
    The order is 1, the reference state is $F_e^{\textrm{max}} = 2$.
  }
  \label{fig:tmp3}
\end{figure}

We test the idea firstly by a one-dimensional model system: one particle in a
tilting double-well potential. For convenience, we let the mass of the
particle to be 1 \textsf{amu}. The unperturbed
Hamiltonian of the system is given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
Notice at room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
See the red line in Fig.~\ref{fig:tmp1}.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect C(\vect p, \vect q) = 0; \qquad
  \vect D(\vect p, \vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$ is  effectively tilt the original
potential $U(\vect q)$. The strength of the perturbation $F_e(t)$ is given
by the following function:
\begin{align}
  F_e(t) = 
  \begin{cases}
    F_e^{\textrm{max}}\times (t / t_c) & t < t_c \\
    F_e^{\textrm{max}} & t \geq t_c
  \end{cases}
\end{align}
$t_c = 20$~\textsf{ps} so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:core-identify-approx-2} is roughly 3~\textsf{ps}.
We choose $\tau = 1$~\textsf{ps}.
% We choose $\mt_\alpha$
% to be the propagator of the Langevin dynamics at temperature $150$
% \textsf{K} with $\alpha = 1$ \textsf{ps}.
Since the dynamics is in
1-d, the projection $\mathcal P$ is identity.  We consider the stability of the
distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
The larger
this value, the more stable the core sets are.

>>>>

\newcommand{\mh}{\mathcal H}
\newcommand{\eps}{\varepsilon}
\newcommand{\ml}{\mathcal L}
\newcommand{\mt}{\mathcal T}
\newcommand{\mo}{\mathcal O}
\newcommand{\mi}{\mathcal I}
\newcommand{\mc}{\mathcal C}
\newcommand{\proj}{\mathit\Pi}
\newcommand{\fwg}{{\mathcal A}}
\newcommand{\bwg}{{\mathcal B}}
\newcommand{\bsigma}{\boldsymbol\sigma}

@@

\title{Linear response theory for core set identification}
\author{Han Wang}
\author{Carsten Hartmann}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin (ZIB), Germany}
% \affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

\section{Theoretical background}
\subsection{Classical equilibrium linear response theory}

In this section, we want to firstly give a brief description of the classical linear response
theory, for details see Ref.~\cite{tuckeman2010statistical} for
example.
We study the non-equilibrium system driven by some external force.
More strictly, we investigate the system that is governed by the
following Langevin equation:
\begin{align}\label{eqn:langevin-1}
  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2}
  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
  + \eps F_e(t) \vect D(\vect q) 
  - \gamma\vect p
  + \sigma\dot{\vect W}.
\end{align}
Here $\vect D$ is the external driving force applied to the system,
the magnitude of which is always assumed to be controllable.
$\eps F_e(t)$ is the strength of the driving as a function of time.
The prefactor $\eps$ indicates that we regard the driving force to be considerably smaller than the internal forces. We will comment on what is meant by "considerably smaller" further below.

% , and
% the magnitude of $F_e(t)$ is small and of order $\mo (\eps)$, $\eps \ll 1$.

Then the phase space distribution $f(\vect x, t)$ ($\vect x = \{\vect
q, \vect p\}$) is subject to the Kolmogorov forward equation:
\begin{align}\label{eqn:orig-forward}
  \frac{\partial}{\partial t} f(\vect x, t) - \fwg^\beta(t) f(\vect x, t) = 0
\end{align}
where $\fwg^\beta(t)$ is the forward infiniesimal generator given by
\begin{align}
  \fwg^\beta(t) =
  \frac{\sigma^2}2\Delta_{\vect p}
  - \nabla_{\vect p}\mh 
  \cdot\nabla_{\vect q}
  - \Big(
  -\nabla_{\vect q}\mh +
  \eps F_e(t)  \vect D  - \gamma\vect p
  \Big)\cdot\nabla_{\vect p}
  + 3N\gamma,
\end{align}
and $\beta2=\gamma / \sigma^2$ is the inverse temperature.
The equilibrium system
is governed by
the standard Langevin equation, i.e. with
vanishing driving force $\eps = 0$ in
Eqn.~\eqref{eqn:langevin-1} and \eqref{eqn:langevin-2}.

Under our smallness assumption on the driving force the full phase space distribution has the following asymptotic expansion in $\eps$:
\begin{align}\label{eqn:f-expan}
  f(\vect x, t) &= f_0(\vect x, t) + \eps f_1(\vect x, t)
  +\eps^2 f_2(\vect x, t) + \mo (\eps^3),
\end{align}
where $f_0=f_0(x,t)$ denotes the evolution of the phase space distribution for the undriven system (with $\eps=0$).
In addition we have:
\begin{align}\label{eqn:A-expan}
  \fwg^\beta(t) = \fwg^\beta_0 + \eps\fwg_1(t),
\end{align}
where
\begin{align}
  \fwg_0^\beta =&
  \frac{\sigma^2}2\Delta_{\vect p}
  -
  \nabla_{\vect p}\mh\cdot\nabla_{\vect q}
  - \big(
  -\nabla_{\vect q}\mh - \gamma\vect p
  \big)\cdot\nabla_{\vect p}
  + 3N\gamma,\\
  \fwg_1(t) =&
  - F_e(t)  \vect D \cdot\nabla_{\vect p}.
\end{align}

The invariant measure of $\fwg^\beta_0$ satisfies $\fwg^\beta_0f^\beta_{\inv}(\vect x) = 0$, and is given by the Boltzmann distribution

@@

  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]

@@

  [\,\fwg^\beta_0 + \eps\fwg_1(t)\,]
  [f_0(\vect x, t) + \eps  f_1(\vect x, t) + \eps^2f_2(\vect x, t)]

@@

Matching this equation at different orders of $\eps$, the zeroth

@@

which is the Kolmogorov forward equation of the undriven system. Matching the first and the second order gives:

@@

  -\beta {F_e(s)}

@@

@@

  -\beta

@@

  - \eps\beta

@@

  + \mo(\eps^2)

@@

Now let $\mathcal{O}$ be an observable of the system. Its expectation value satisfies the linear response formula:

@@

  [\,f_{\inv}^\beta(\vect x) + \eps f_1(\vect x, t)] +\mathcal{O}(\eps^2)\\\nonumber
   \label{eqn:eqi-pert-1}

@@

 \eps \beta

@@

  O(\vect x)+\mathcal{O}(\eps^2)

@@

where $ \langle O\rangle=\int \dd\vect x\,O(\vect x) f_{\inv}^\beta(\vect x)$ is the expectation value of the undriven equilibrium system, and $\bwg^\beta_0$ is the infiniesimal generator of the undriven 
Kolmogorov backward equation, which is the adjoint operator of $\fwg^\beta_0$. That is,
we have

@@

where $\vect X^\beta_t$ is the trajectory of the undriven Langevin equation (with $\eps=0$)

@@

Therefore, we have up to order 2 in $\eps$:

@@

 & =

@@

  \eps\beta

@@

  \mathbb E_{\vect x} [O(\vect X^\beta_s)]\\
\label{eqn:eq-lr}  
 & =

@@

  \eps \beta

@@

  j\cdot
  \mathbb E [O(\vect X^\beta_s)]

@@

  correlation} between the observation $O$ at time $s$ and the flux $j$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The classical response theory mainly has the following two limitations:
Firstly, one has to start from an equilibrium simulation. That means if the
external driving is no longer small ($\eps$ is relatively big), the
first order (or linear) response~\eqref{eqn:eq-lr} may not be a good approximation to
the true non-equilibrium process any more.
One possible way to improve the accuracy is to introduce the second
order response term, e.g., based on equation~\eqref{eqn:f2-0}.
Here one encounters the second limitation of the classical
response theory: it does not provide a frame work, under which
the second and even higher order responses can be easily derived.
The limitations of the classical responses theory motivate us to
develop a more general framework that allows one starting from any
non-equilibrium process (arbitrarily driven), and calculating linear responses from it.

This time, we consider the following Langevin equation:
\begin{align}\label{eqn:langevin-1b}
  \dot{\vect q} & = \nabla_{\vect p}\mh(\vect q,\vect p), \\\label{eqn:langevin-2b}
  \dot{\vect p} & =- \nabla_{\vect q}\mh(\vect q,\vect p)
  + (k_0 F_e(t)+\eps \Delta F_e(t)) \vect D(\vect q)
  - \gamma\vect p
  + \sigma\dot{\vect W},
\end{align}
such that for $\eps=0$ we get a driven system that is further
perturbed by the addition time-dependent driving force $\eps \Delta
F_e(t) \vect D(\vect q)$. Let $\vect X^\eps_t$ denote the associated solution process, and
$\mathcal P_\eps$ the probability measure of the trajectories
generated by driven Langevin
dynamics~\eqref{eqn:langevin-1b}--\eqref{eqn:langevin-2b}.

The time-dependent expectation value $O_0(t)$ of the path observable $O$ under the dynamics of the Langevin equation (\ref{eqn:langevin-1b})-(\ref{eqn:langevin-2b}) for $\eps=0$
can be written as
\begin{align}
  O_0( t) =
  \langle O[\vect X^0_s]\,\rangle_{0,t} = 
  \int\dd \vect u\, f(\vect u, 0) 
  \int_{\mc\{\vect u,0; t\}}
  O[\vect X_s]\, \dd\mathcal P_0[\vect X_s],
\end{align}
where 
$\langle \cdot \,\rangle_{0,t}$ is the ensemble average over all possible trajectories for $\eps=0$ in time interval $[0,t]$,
$f(\vect x, 0)$ is the initial distribution that may not be the equilibrium distribution, and
$\mc\{\vect u,0; t\}$ is the set of all admissible continuous trajectories starting from point
$\vect u$ at time $0$, and ending at some position at time $t$.

For $\eps\ge 0$ the  expectation value $O_\eps(t)$ takes the form
\begin{align}
  O_\eps(t) & =
  \int \dd\vect u f(\vect x,0)
  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
  \, \dd\mathcal P_\eps[\vect X_s]   \\
& = \int \dd\vect u f(\vect x,0)
  \int_{\mc\{\vect u,0; t\}} O[\vect X_s]\,
  \frac{\dd {\mathcal P}_\eps[\vect X_s]}{\dd\mathcal P_0[\vect X_s]}
  \, \dd\mathcal P_0[\vect X_s] .   
\end{align}
By using the Girsanov formula, 
\begin{align}\label{eqn:girsanov}
  \frac
  {\dd{\mathcal P_\eps}[\vect X_s]}
  {\dd\mathcal P_0[\vect X_s]}  
  =&\,
  \exp
  \big\{{\eps G[\vect X_s] - \eps^2 H[\vect X_s]}
  \big\}
\end{align}
where the short notations are defined by:
\begin{align}
  G[\vect X_s]
  &= \frac{1}\sigma\int_0^t
  \Delta F_e(s)\,\vect D(\vect q_s)\cdot\dd\vect W_s \\
  H[\vect X_s]
  &=
  \frac 1{2\sigma^2}\int_0^t
  \Delta F_e^2(s)\,\vect D^2(\vect q_s)\,\dd s.
\end{align}
Expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov} in $\eps$, we have
\begin{align}\label{eqn:neq-response}
   O_\eps(t) -
  O_0(t)
  =
  \eps\, \big\langle O[\vect X_s]\cdot G[\vect X_s]\big\rangle_{0,t}
  +
  \frac{\eps^2}{2} \big\langle O[\vect X_s] \cdot (G^2[\vect X_s] - H[\vect X_s] )\big\rangle_{0,t}
  +
  \mo (\eps^3)
\end{align}
Some remarks concerning Eq.~\eqref{eqn:neq-response} seem in order:
\begin{itemize}
\item The first term on the right-hand-side of Eq.~\eqref{eqn:neq-response} is the linear
  response to the reference non-equilibrium process (driven by $F_e$, $\eps=0$). The next term is the second order response term. Obviously, one can easily write down higher orders of response by the Taylor expansion of the right-hand-side of Eq.~\eqref{eqn:girsanov}.
\item When the observable $O[\vect X_s]$ only depends on the end point of the trajectory, i.e.
  $O[\vect X_s] = O(\vect X_t)$, then $\big\langle O(\vect X_t)\cdot G[\vect X_s]\big\rangle_{0,t}$
  gives the well-know time correlation form of the linear response formula.
\item If one wants to calculate the same expectation value for a family of non-equilibrium perturbations $\Delta F_e(t)$ then 
one does not need to repeat the calculations of Eq.~\eqref{eqn:neq-response} for every member of the family. 
  If it is possible to
  express different $\Delta F_e$ in the same basis, then the responses must only be
  calculated for the single basis functions. Then with a linear combination of the responses on basis functions, one can derive
  the responses of the whole family. 
\item In the above derivation, we assume that the reference and the perturbed non-equilibrium processes
  have the same initial distribution (both of them may be different from the equilibrium distribution),
  but this is not a necessary. One can easily
  apply a similar reweighing approach to the initial distribution as we used it for the trajectory ensemble.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Numerical tests: the core set identification of  the Markov State Model }

The core set plays a very important role in the MSM theory.  We
therefore want to test our idea of non-equilibrium response theory by
identifying the core sets in a non-equilibrium driven system.
There are several ways to define a core set.
One possible choice  is to calculate the stability of the phase space
distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
x, t)$. And the core sets are defined by
\begin{align}
  C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
\end{align}
Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
frozen, and the distribution $f$ evolves a long the artificial time
$\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
be chosen smaller than the investigated relaxation time scales of the
original system. However, on the other hand, $\tau$ cannot be
arbitrarily small, because otherwise the statistical uncertainty would
not allow us to distinguish $\mt_\tau f$ and $f$.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
  \caption{The double-well potential with tilting perturbation at
    time $t\geq 20$ ps..}
  \label{fig:tmp1}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str2-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$ under
    perturbation with strength $\eps = 2$.  From left to right the
    columns present results at time $t = 0$, 10, 15 and
    20~\textsf{ps}.  The 1st row: the direct non-equilibrium
    simulation. The 2nd row: the results of traditional equilibrium
    linear response theory.  The 3rd row: the results of the response
    theory developed by this work.  We calculate response terms up to
    the second order from an equilibrium simulation (i.e. letting $k_0
    = 0$ and $\eps = 0$ in Eq.~\eqref{eqn:langevin-2b}).
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str4-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $\eps = 4$. From left to right the
    columns present results at time $t = 0$, 10, 15 and
    20~\textsf{ps}.
    The first three rows use the same algorithm as Fig.~\ref{fig:tmp2}.    
    The 4th row: the results of the response theory developed by this work.
    Instead of calculating the responses from the equilibrium simulation,
    here we calculate the linear response from a non-equilibrium simulation
    of strength 2 (i.e. $k_0 = 2$ and $\eps = 4$
    in Eq.~\eqref{eqn:langevin-2b}).
    Therefore, the effective perturbation in this case is only
    $\eps = 2$.
  }
  \label{fig:tmp3}
\end{figure}

We test the idea firstly by a one-dimensional model system: one particle in a
tilting double-well potential. For convenience, we let the mass of the
particle to be 1~\textsf{amu},
and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
The unperturbed
Hamiltonian of the system is given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$   effectively tilts the original
potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
The strength of $F_e(t)$ is
set to be 1, when the time $t$ is larger than some warm-up time $t_c$.
When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
i.e. $ F_e(t) = t/t_c$.
We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
We choose $\tau = 1$~\textsf{ps}, and
% We choose $\mt_\alpha$
% to be the propagator of the Langevin dynamics at temperature $150$
% \textsf{K} with $\alpha = 1$ \textsf{ps}.
consider the stability of the
distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
The larger
this value, the more stable the core sets are.

See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
the results of the brute-force non-equilibrium simulation
and that of the response theory are presented.
In Fig.~\ref{fig:tmp2},
the good agreement between the brute-force (1st row)
and the linear response result (2nd row) indicates that $\eps = 2$ is
indeed a relatively small perturbation to the equilibrium.
In the third row of Fig.~\ref{fig:tmp2},
the second order response is considered. We do not see any
substantial improvement of the result by including this higher order term,
because the
driving force to the system is small,
and the linear term is anyway a good approximation.

In Fig.~\ref{fig:tmp3}, as the
perturbation grows stronger, the left well becomes less stable, while
the right well becomes more stable.  When the perturbation is as strong as
$\eps = 4$, the linear response result (2nd row) is no
longer precise:
At $t\geq 20$ \textsf{ps}, the left well basically
disappears, however, the linear response calculation presents
artificial structure (that is a small artificial
stable region accompanied by a small unstable region)
at the position of the left well.
The second order response (3rd row) improves the accuracy a little, 
however, it also includes artificial finer-grained oscillation in the profile.
At the same time, the numerical uncertainty is higher, so the figure
looks noisy.
In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
that calculate from 
a intermediate
non-equilibrium simulation with driving force of $k_0 = 2$.
Notice that the reference simulation in this case is also a
non-equilibrium simulation, so the effective strength of perturbation
is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
considered as a small perturbation.
A substantial improvement is achieved: the profile
is correctly calculated, and the statistical error is small.
